---
title: "Geography 176A"
author: "[Ryan Erickson](https://github.com/IamRyanE)"
subtitle: 'Lab 04: Distances and the Border Zone'
output:
  html_document:
    theme: journal
---

# Libraries

### Functions
```{r}
source("../R/utils.R")
```

### SPDS
```{r, message=FALSE}
library(tidyverse)
library(sf)
library(units)
library(leaflet)
```

### Data
```{r, message=FALSE}
library(USAboundaries)
library(rnaturalearth)
library(readxl)
```

### Visualization
```{r, message=FALSE}
library(gghighlight)
library(ggrepel)
library(ggthemes)
library(knitr)
library(kableExtra)
library(rmapshaper)
```

## Question 1: Tesselation of Surfaces

#### Step 1.1
Getting CONUS States (without function)
```{r}
conus = USAboundaries::us_counties() %>% 
  filter(!state_name %in% c("Hawaii", "Puerto Rico", "Alaska",
             "Guam", "District of Columbia")) %>% 
  st_transform(5070)
```
 
#### Step 1.2
```{r}
conus_cent = conus %>% 
  st_centroid() %>% 
  st_combine()

conus_un = conus %>% 
  st_union() %>% 
  ms_simplify(keep = .05)
```

#### Step 1.3 & Step 1.4 & 1.5
```{r}
conus_vor = conus_cent %>% #1.3
  st_voronoi() %>% 
  st_cast() %>% 
  st_as_sf() %>%
  mutate(id = 1:n()) %>% 
  st_intersection(conus_un)
  
conus_tri = conus_cent %>% 
  st_triangulate() %>% 
  st_cast() %>% 
  st_as_sf() %>%
  mutate(id = 1:n()) %>% 
  st_intersection(conus_un)

conus_grd = conus_cent %>% 
  st_make_grid(n = c(70,45)) %>% 
  st_cast() %>% 
  st_as_sf() %>%
  mutate(id = 1:n()) %>% 
  st_intersection(conus_un)

conus_hex = conus_cent %>% 
  st_make_grid(n = c(70,50), square = FALSE) %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  mutate(id = 1:n()) %>% 
  st_intersection(conus_un)
```

#### Step 1.6 - FUNCTION JUNCTION
```{r}
plot_tes = function(arg1, title){
  ggplot() + 
    geom_sf(data = arg1, col = "white", size = .2) + 
    theme_void() + 
    theme(legend.position = 'none',
          plot.title = element_text(face = "bold", color = "darkblue", hjust = .5, size = 24)) +
    labs(title = paste0(title),
         caption = paste0(nrow(arg1), " features"))
}
```

#### Step 1.7
```{r}
plot_tes(conus, "Original")
plot_tes(conus_vor, "Voroni")
plot_tes(conus_tri, "Triangualtion")
plot_tes(conus_grd, "Square")
plot_tes(conus_hex, "Hexegonal")
```

# Question 2
#### Step 2.1 - FUNCTION JUNCTION PART 2
```{r}
sum_tes = function(arg1, string){
  
  newframe = arg1 %>% 
    mutate(area = st_area(arg1)) %>% 
    drop_units() %>% 
    mutate(area = area*0.000001)

  dataframe = data.frame(type = string,
                num_feat = nrow(newframe),
                mean_area = mean(newframe$area),
                st_dev = sd(newframe$area),
                tot_area = sum(newframe$area))
 
  return(dataframe)
}
```

#### Step 2.2
```{r}
sum_tes(conus, "original")
sum_tes(conus_vor, "voronoi")
sum_tes(conus_tri, "triangualtion")
sum_tes(conus_grd, "square grid")
sum_tes(conus_hex, "hexegonal")
```

#### Step 2.3
```{r}
tess_summary = bind_rows(
  sum_tes(conus, "original"),
  sum_tes(conus_vor, "voronoi"),
  sum_tes(conus_tri, "triangualtion"),
  sum_tes(conus_grd, "square grid"),
  sum_tes(conus_hex, "hexegonal"))
```

#### Step 2.4
```{r}
kable(tess_summary,
             col.names = c("Type", "Number of Features", "Mean Area", "St. Dev", "Total Area"),
             caption = "Summary of Our 5 Tesselations") %>% 
           kable_styling(bootstrap_options = "striped", full_width = F)
```

#### Step 2.5
Describe the differences in each tessellation (original, voronoi, triangulation, etc). Which tessellation would be best for a MAUP? Which tessellation would be the easiest to compute? If you wanted to accurately represent the US which would you choose?  

# Question 3
#### Step 3.1
```{r, message = FALSE, results = FALSE, warning=FALSE}
dam_data = read_excel("../data/NID2019_U.xlsx") %>% 
  filter(!is.na(LATITUDE)) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) %>% 
  st_transform(5070) %>% 
  st_filter(conus_un)
```

#### Step 3.2
```{r}
point_in_polygon = function(points, polygon, nam){
  st_join(polygon, points) %>% 
    st_drop_geometry() %>% 
    count(get(nam)) %>% 
    setNames(c(nam, "n")) %>% 
    left_join(polygon, by = nam) %>% 
    st_as_sf()
}
```

```{r}
con_pip = point_in_polygon(dam_data, conus, "geoid")
vor_pip = point_in_polygon(dam_data, conus_vor, "id")
tri_pip = point_in_polygon(dam_data, conus_tri, "id")
grd_pip = point_in_polygon(dam_data, conus_grd, "id")
hex_pip = point_in_polygon(dam_data, conus_hex, "id")
```

```{r}
plot_counts = function(arg1, title){
 
   ggplot() + 
    geom_sf(data = arg1, aes(fill = n), col = NA, size = .2) + 
    viridis::scale_color_viridis() + 
    theme_void() + 
    theme(plot.title = element_text(face = "bold", color = "darkblue", hjust = .5, size = 24)) +
    labs(title = paste0(title),
         caption = paste0("Total number of dams present: ", sum(arg1$n)))
}
```

#### Step 3.6
```{r}
plot_counts(con_pip, "Dams in US: Originial")
plot_counts(vor_pip, "Dams in US: Voronoi")
plot_counts(tri_pip, "Dams in US: Triangulation")
plot_counts(grd_pip, "Dams in US: Square Grid")
plot_counts(hex_pip, "Dams in US: Hexagonal Grid")
```

# Question 4
#### Step 4.1
```{r}
unique(dam_data$PURPOSES) %>% length

dam_data$PURPOSES[1:5]

nid_classifier = data.frame(abbr = c("I","H","C","N","S","R","P","F","D","T","G","O"),
                            purpose = c("Irrigation","Hydroelectric","Flood Control","Navigation","Water Supply","Recreation","Fire Protection","Fish and Wildlife","Debris Control","Tailings","Grade Stabilization","Other"))


dam_freq <- strsplit(dam_data$PURPOSES, split = "") %>%
  unlist() %>% 
  table() %>% 
  as.data.frame() %>% 
  setNames(c("abbr", "count")) %>% 
  left_join(nid_classifier) %>% 
  mutate(lab = paste0(purpose, "\n(", abbr, ")"))
```
The above is just test code to see what Mike was talking about and to clarify more about the types of dams. Some dams can have multiple types, i.e a dam of FR type is a recreational type dam that is either protected by or used by the Fish and Wildlife Service. 

Here we're going to choose our dams
```{r}
majorrivers = read_sf("../data/MajorRivers.dbf") %>% 
  filter(SYSTEM == "Mississippi") %>%
  mutate(STATE = c("AR", "MI", "MO", "OH")) %>% 
  st_transform(4326)

dams_of_interest = dam_data %>% 
  filter(grepl("H", PURPOSES)) %>%
  filter(DAM_NAME != "SOO LOCKS") %>% 
  select(DAM_NAME, PURPOSES, NID_STORAGE, YEAR_COMPLETED, STATE) %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs=4326) %>% 
  st_transform(st_crs(majorrivers)) %>% 
  group_by(STATE) %>% 
  slice_max(NID_STORAGE, n=1) %>% 
  ungroup()

leaflet() %>% 
  addProviderTiles(providers$CartoDB) %>% 
  addPolylines(data = majorrivers) %>% 
  addCircleMarkers(data = dams_of_interest,
             fillColor  = ~colorQuantile("YlOrRd", NID_STORAGE)(NID_STORAGE),
             color = NA,
             fillOpacity = .5,
             radius = ~NID_STORAGE / 1500000,
             label = ~DAM_NAME,
             popup = leafpop::popupTable(st_drop_geometry(dams_of_interest), feature.id = FALSE, row.numbers = FALSE)) %>% 
  addMeasure() %>% 
  leafem::addMouseCoordinates() %>% 
  leafem::addHomeButton(group = "Missouri")
```

